---
title: "Machine Learning Project"
author: "Nizamo"
date: "Sunday, January 31, 2016"
output: word_document
---
Prediction Assignment Writeup - Practical Machine Learning
==========================================================
#Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project is come from this source: http://groupware.les.inf.puc-rio.br/har. This dataset is licensed under the Creative Commons license (CC BY-SA).Please visit the site for further explaination of this marvellous group.

#Objective
The objective of this project is to predict the manner of performing unilateral dumbbell biceps curls based on data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. The 5 possible methods include -
A: exactly according to the specification
B: throwing the elbows to the front
C: lifting the dumbbell only halfway
D: lowering the dumbbell only halfway
E: throwing the hips to the front

##############################################################################
1. Downloaded the data and save it in your working directory. Check the default directory and change it to your working directory.
getwd()
setwd("E:/coursera nizam/Course Project Module 8")
```{r}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
set.seed(12345)
options(warn=-1)
```
Values contained "#DIV/0!" have been replaced by an NA value.
```{r}
training_data <- read.csv("E:/coursera nizam/Course Project Module 8/pml-training.csv", na.strings=c("#DIV/0!") )
testing_data <-  read.csv("E:/coursera nizam/Course Project Module 8/pml-testing.csv", na.strings=c("#DIV/0!") )
```
All columns 8 to the end have been casted to numeric.
```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}
for(i in c(8:ncol(testing_data)-1)) {testing_data[,i] = as.numeric(as.character(testing_data[,i]))}
```
A good prediction cannot be achieved as some of the columns mostly blank.  Feature sets that include complete column only will be chosen. Username, timestamps and windows also have been removed. Determination of feature sets have been done and displayed.  
```{r}
feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```
Model data built from selected feature set have been obtained.
```{r}
index <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[index,]
testing <- model_data[-index,]
```
Five random forests with 150 trees each have been developed and parallel processing have been used for building this model. 
```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe
rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```
Error reports for both training and test data have been generated.
```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)
predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

Conclusions and Test Data Submit
--------------------------------

Confusion matrix obtained from this model is very accurate. Because test data was around 99% accurate, it is expected nearly all of the submitted test cases to be correct.  It turned out they were all correct.


```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

x <- testing_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)
answers

```
